\documentclass[letterpaper]{article}  % list options between brackets
\usepackage{authblk,graphicx,url}              % list packages between braces
% type user-defined commands here

\renewcommand\Affilfont{\footnotesize}

\begin{document}

\title{ARA Web Plotter \\ \small{Based on QNDWBOM -- Quick 'n' Dirty Web-Based Online Monitoring} \\ \small{Version : SVN 315 or 1.3}}   
\author{R.~Nichol}

\affil{Dept. of Physics and Astronomy, UCL, Gower Street, London, WC1E 6BT, UK}

\maketitle


\begin{abstract}
  This is intended to be a quick-start and trouble shooting guide for running the ARA Web Plotter code. The code is currently under development and desired features can be requested by contacting the author.
\end{abstract}

\section{Introduction}
The ARA web plotter code is designed to analyse raw data from the ARA TestBed1 producing plots which are viewable on a website. Most of these plots show the time dependence of variables from the event and housekeeping data structures.

\section{Components}
The ARA web plotter code consists of a number if quasi-independent programs:
\begin{itemize}
\item Script to convert raw data into ROOT data (runAraRunFileMaker.sh)
\item AraWebPlotter the program that reads the ROOT format data and produces summary plots for each run
\item AraTimeWebPlotter the program that reads the output files of AraWebPlotter to produce plots over a longer timespan than a single run
\item getDirFromConfig a utility program to read the config file and return the lcoation of a specified working directory (see below)
\end{itemize}


\section{Working Directories}\label{workdir}
There are two directories used by the ARA web plotter for managing the ROOT data and output plots. These directories are configured using the araWebPlotter.config file. The araWebPlotter.config file is located using the {\tt ARA\_WEBPLOTTER\_CONFIG\_DIR} environmental variable.
\begin{itemize}
\item plotDir -- The web visible directory to which plots are written
\item rootFileDir -- The directory used for storing the ROOT data files used by the web plotter
\end{itemize}

\section{Installation}
\subsection{Prerequisites}
The ARA ROOT software (of which the web plotter is part) depends on ROOT and a number of other libraries. These are summarised below:
\begin{itemize}
\item FFTW3 -- \url{http://www.fftw.org}
\item Gnu Scientific Library -- \url{http://www.gnu.org/software/gsl/}
\item ROOT -- \url{http://root.cern.ch} (it is important that the MathMore package gets built)
\item libRootFftwWrapper -- \url{http://www.hep.ucl.ac.uk/uhen/libRootFftwWrapper}
\end{itemize}

\subsection{Compilation}
Once these items are installed (in the order above is easiest), the AraRoot libraries can be compiled. The recommended method for compilation is to set the {\tt ARA\_UTIL\_INSTALL\_DIR} environmental variable to point to the desired installation directory, the isntallation directory will default to /usr/local if this is not set. Compilation should be as simple as typing \begin{verbatim}make\end{verbatim} in the AraRoot directory.

\subsection{installation}
To install the libraries, header files and executables type \begin{verbatim}make install\end{verbatim} in the AraRoot directory. You need ensure that {\tt ARA\_UTIL\_INSTALL\_DIR/lib} is in your LD\_LIBRARY\_PATH and {\tt ARA\_UTIL\_INSTALL\_DIR/bin} is in your PATH.

To install the html files associated with the ARA Web Plotter the recommended method is:
\begin{enumerate}
\item Ensure that the ARA\_WEBPLOTTER\_CONFIG\_DIR environmental variable is set and points to the directory containing araWebPlotter.config
\item Ensure that araWebPlotter.config contains the correct paths for the four working directories
\item Check that the \begin{verbatim}getConfigDir --plotDir\end{verbatim} returns the directory in which the html files are to be installed
\item Edit the AraWebPlotter/installHtmlFiles.sh to set the appropriate HTML\_REL\_DIR for the local system (i.e. if the website for the ARA WebPlotter is visible at \url{http://www.hep.ucl.ac.uk/uhen/ara/monitor/}, the the HTML\_REL\_DIR shoudl be set to /uhen/ara/monitor
\item Change to the AraWebPlotter directory and execute ./installHtmlFiles.sh
\end{enumerate}

Ensure the other working directories described in Section~\ref{workdir} exist and have read/write permissions set correctly.


\section{Running Web Plotter}
If everything is set up correctly running the web plotter should involve these steps:
\begin{enumerate}
\item For each new run create a ROOT version of the data using runAraRunFileMaker.sh. In order to do this one has to define a directory on the local system to store the ROOT data and edit the runAraRunFileMaker.sh to point to the correct directories for input (raw) and output (root) data. Running the script when these are set is a simple as \begin{verbatim} runAraRunFileMaker.sh <run no>\end{verbatim}
\item For each new ROOT file created in step 1, process the data with the web plotter. If there are both event and hk files both can be processed otherwise the event file on it's own can be processed. To run the web plotter one does: \begin{verbatim} AraWebPlotter /path/to/the/event/file/eventXXX.root /path/to/the/hk/file/hkXXX.root \end{verbatim}
\item Periodically one can update the plots that show the behaviour across multiple runs. To make these time plots: \begin{verbatim} AraTimeWebPlotter \end{verbatim}
\end{enumerate}


\section{Remote ROOT Access}
It is possible to remotely interactively access the ROOT data using the web interface. Although this is only useful for very quick checks for most analysis needs it will be more efficient to download the files and store them locally. Provding that the directory containing the root subdirectories is hosted in the webserver. For instance one could do \\ \begin{verbatim}TFile::Open("http://www.hep.ucl.ac.uk/uhen/ara/monitor/root/run1154/event1154.root")\end{verbatim} or by starting AraDisplay with the event file string as \begin{verbatim}"http://www.hep.ucl.ac.uk/uhen/ara/monitor/root/run1154/event1154.root"\end{verbatim}. 

As similar thing could be done using the data hosted on the web servers at Wisconin by passing the http username and password in the string, e.g. \begin{verbatim}''http://username:password@convey.icecube.wisc.edu/data/exp/ARA/2011/filtered/L0/0205/run001669.root''\end{verbatim}.

% \begin{itemize}
% \item The run files which contain only events (and housekeeping) from given runs.
% \item The complete backup files which contain all data of that type from the flight. These backup files are written once and hour and are tagged 1 and 2 for historical reasons. The most recent file supercedes all previous files. So the files from December 26th contain all the data in the files from December 25th.
% \end{itemize}

\section{Event Plots}
There is an optional flag in the config file that governs whether or not plots of all events are made. If these plots are made there is a substantial slow down of the webplotter code. Currently on my desktop machine at UCL it takes of order 16 seconds to process a run without event plots, and around factor of ten slower with the event plotting turned on. It isn't clear to me how useful the event plots are, but opinions may vary.

\section{Current Status}
The UCL version of the webplotter is hosted at \url{http://www.hep.ucl.ac.uk/uhen/ara/monitor/}

At the time of writing there are several omissions from the current web plotter. The planned upgrades (but possibly with an infinite timescale) include:
\begin{itemize}
\item{Snapshot page showing last housekeeping info (rather than just the time dependence}
\item{Anything else that the users request}
\end{itemize}

\section{Troubleshooting}
This section will be updated as information becomes available.

\newpage

\section{Config file  araWebPloter.config}
\begin{verbatim}
<output>
rootFileDir#S=/unix/anita1/ara/data/webplotter;
plotDir#S=/unix/www/html/uhen/ara/monitor;
makeEventPlots#I1=0;
</output>
\end{verbatim}

 
\end{document}

